Testing traversability of protein tunnels with flexible ligands using motion planning

### 2017-06-14 paper.tex
table  Please enter the commit message for your changes. Lines starting

\def\L{L}

@@

To guide thee growth of the tree through the tunnel, a moving virtual goal is used~\cite{vonasek2009rrt}.
The virtual goal $v$ is the index of a sphere, around which the random samples are generated with probability $\gb$.
%The random samples are not generated only from whole $\C$, as in the original RRT, but the sampling
%is guided using moving virtual goal~\cite{vonasek2009rrt}.
After the tree reaches the sphere $t_v \in T$, i.e., the distance of the tree to the center of $t_v$ is
less than a predefined threshold $\dt$, the virtual goal is moved to successor of the last sphere in the tunnel
that is reached by the tree (lines~\ref{alg::main:a}--\ref{alg::main:b} in Alg.~\ref{alg::main}).
Setting the virtual goal to this successor allows the tree to ``skip'' such parts of the tunnels that are
not traversible or reachable by the ligand.
The algorithm terminates after a predefined planning trials $\Imax$ or if the tree reaches
the last sphere in the tunnel, i.e., when $v = N$.

@@

\linesnumbered
\begin{algorithm}[h]
%\setstretch{\straa}
\caption{\label{alg::main}main loop of the RRT planner}
\KwIn{
    tunnel $T=(t_i, r_i), n=1,\ldots N$ represented by $N$ spheres,
    initial configuration $\qinit$,
    goal configuration $\qgoal$,
}
\KwData{
   ligand conformations $\L$,
   scale limits $\smin, \smax$ and $\sdelta$
}
\KwOut{
    configuration tree $\T$\;
}
\hrule
$v = 0$; // index of the virtual goal\\
$iteration = 0$\;
\While{$iteration < \Imax$ {\bf and}  $v < N$}{
    \eIf{$rand() < \gb$}{
        $\qrand$ = random sample around actual virtual goal $t_n \in T$\;
    }{
        $\qrand$ = random sample from $\C$\;
    }
    $\qnear$ = nearest node in $\T$ towards $\qrand$\;

@@

    \For{$i= N,N-1,\ldots v+1,v$}{ \nllabel{alg::main:a}
        $d$ = nearest node in the tree to $T[i]$\;
        \If{$d < \dt $}{
            $v = i+1$; // new virtual goal found\;
            {\bf  break}\;
        }
    } \nllabel{alg::main:b}
    $iteration = iteration+1$\;
}
\return $\T$\;
\end{algorithm}

The core of the proposed method is the expansion procedure (Alg.~\ref{alg::expand}) which generates new collision-free nodes around $\qnear\in\T$.
For each ligand conformation, the expansion procedure attempts to find a new collision-free configuration around $\qnear$ with maximal
scale.
For a given conformation $l \in \L$, first the maximal scale $\smax$ is considered and
$m$ samples are generated around $\qrand$ and tested for collision.
The nearest collision-free sample towards $\qrand$ is selected and added to the tree.
If none of the samples is collision-free, the scale is reduced to $\smax-\sdelta$ and the searches continues
until a free collision-free samples is found or until the mimimal reduced-scale $\smin$ is reached.

@@

>>>>

\def\L{L}
\def\LA{L_1}
\def\LB{L_2}

@@

To guide thee growth of the tree through the tunnel, a moving virtual goal is used~\cite{vonasek2009rrt}.
The virtual goal $v$ is the index of a sphere, around which the random samples are generated with probability $\gb$.
%The random samples are not generated only from whole $\C$, as in the original RRT, but the sampling
%is guided using moving virtual goal~\cite{vonasek2009rrt}.
After the tree reaches the sphere $t_v \in T$, i.e., the distance of the tree to the center of $t_v$ is
less than a predefined threshold $\dt$, the virtual goal is moved to successor of the last sphere in the tunnel
that is reached by the tree (lines~\ref{alg::main:a}--\ref{alg::main:b} in Alg.~\ref{alg::main}).
Setting the virtual goal to this successor allows the tree to ``skip'' such parts of the tunnels that are
not traversible or reachable by the ligand.
The algorithm terminates after the predefined number of planning trials $\Imax$ or if the tree reaches
the last sphere in the tunnel, i.e., when $v = N$.

@@

\linesnumbered
\begin{algorithm}[h]
%\setstretch{\straa}
\caption{\label{alg::main}main loop of the RRT planner}
\KwIn{
    tunnel $T=(t_i, r_i), n=1,\ldots,N$ represented by $N$ spheres,
    initial configuration $\qinit$,
    goal configuration $\qgoal$,
}
\KwData{
   ligand conformations $\L$,
   scale limits $\smin, \smax$ and $\sdelta$
}
\KwOut{
    configuration tree $\T$\;
}
\hrule
$v = 0$; // index of the virtual goal\\
$iteration = 0$\;
\While{$iteration < \Imax$ {\bf and}  $v < N$}{
    \eIf{$rand() < \gb$}{
        $\qrand$ = random sample around actual virtual goal $t_n \in T$\;
    }{
        $\qrand$ = random sample from $\C$\;
    }
    $\qnear$ = nearest node in $\T$ towards $\qrand$\;

@@

    \For{$i= N-1,N-2,\ldots,v+1,v$}{ \nllabel{alg::main:a}
        $d$ = nearest node in the tree towards sphere $t_i$\;
        \If{$d < \dt $}{
            $v = i+1$; // new virtual goal found\;
            {\bf  break}\;
        }
    } \nllabel{alg::main:b}
    $iteration = iteration+1$\;
}
\return $\T$\;
\end{algorithm}

The core of the proposed method is the expansion procedure (Alg.~\ref{alg::expand}) which generates new collision-free nodes around $\qnear\in\T$.
For each ligand conformation $l \in \L$, the expansion procedure attempts to find a new collision-free configuration around $\qnear$ with a maximal
scale.
First, the maximal scale $\smax$ is tested and $m$ random samples are generated around $\qrand$ and tested for collision.
The nearest collision-free sample towards $\qrand$ is selected and added to the tree.
If none of the tested samples is collision-free, the scale is reduced to $\smax-\sdelta$ and the searches continues
until a free collision-free samples is found or until the mimimal reduced-scale $\smin$ is reached.

@@

The proposed method has been used to analyze a tunnel in Haloalkane dehalogenase (protein 1CQW) (Fig.~\ref{fig::tunnel}).
The tunnel was computed using CAVER 3.0~\cite{caver3} for spherical probe of radius 0.9~\AA.
The tunnel is represented by 760 spheres with average distance $0.05$~\AA.

% from David: m003 je "1-chlorpropan" a m004 "1-chlorbutan"
The traversability of the tunnel was evaluated for 1-chlorpropan (denoted as $\LA$) and 1-chlorbutan (denoted as $\LB$).

\begin{figure}
\centering
\caption{\label{fig::tunnel}
    TODO image of the tunnel + 1cqw
}
\end{figure}

\begin{table}
\centering
\caption{\label{tab::rrt}
    Runtime and sucess ratio of the planner
}
\input table.003.tex
\hskip 10pt
\input table.004.tex
\end{table}

Further information can be found at {\url{http://mrs.felk.cvut.cz/isrr2017}}.
